% iaus2esa.tex -- sample pages for Proceedings IAU Symposium document class
% (based on v1.0 cca2esam.tex)
% v1.04 released 17 May 2004 by TechBooks
%% small changes and additions made by KAvdH/IAU 4 June 2004
% Copyright (2004) International Astronomical Union

\NeedsTeXFormat{LaTeX2e}

\documentclass{iau_FM}
\usepackage{graphicx}
\usepackage{hyperref}

\title[Andean ROAD] %% give here short title %%
{The new Andean Regional Office \\
of Astronomy for Development}

\author[Farid Char \& Jaime Forero-Romero]   %% give here short author list %%
{Farid Char$^1$
%%  \thanks{Present address: Fluid Mech Inc., 24 The Street, Lagos, Nigeria.},
 \and Jaime Forero-Romero$^2$}

\affiliation{$^1$Unidad de Astronom\'ia, Facultad de Cs. B\'asicas, Universidad de Antofagasta \\ Avenida U. de Antofagasta 02800, Antofagasta 1270300, Chile \\ tel.: +56 55 2637596, email: {\tt farid.char@uantof.cl} \\[\affilskip]
$^2$Departamento de F\'isica, Universidad de los Andes \\ Carrera 1 18-10, Bloque Ip., Bogot\'a, Colombia \\ tel.: +571 3394949-5183, email: {\tt je.forero@uniandes.edu.co}}

\pubyear{2015}
\setcounter{page}{1}
\jname{Astronomy in Focus, Volume 1} 
\editors{Piero Benvenuti, ed.}
\begin{document}

\maketitle

\begin{abstract}
The Andean Regional Office of Astronomy for Development (ROAD) is a
new effort in South America to serve several goals in astronomical
development. The six countries in the Andean ROAD (Bolivia, Colombia,
Chile, Ecuador, Peru and Venezuela) represent a common language block
in the region. They work together to develop strategies to strengthen the
professional research, education and popularization of astronomy. Our
current Working Structure comprises a ROAD Coordinator and one
Coordinators in each Task Force. Here we describe the main points of
the ROAD's current action plan.
 
\keywords{andean, office, development}
%% add here a maximum of 10 keywords, to be taken form the file <Keywords.txt>
\end{abstract}

\firstsection % if your document starts with a section,
              % remove some space above using this command.
\section{Introduction}

The countries in the Andean Region (Bolivia, Colombia, Chile, Ecuador,
Peru and Venezuela) represent a common language block in South
America, sharing a common vision about the importance of cultural
development that can be achieved through regional cooperation and
institutional commitment. 
The Andean Regional Office of Astronomy for Development (ROAD) serves
this goal strengthening ongoing collaboration efforts, creating
channels of communication and developing new strategies to exchange
knowledge and human resources. 
The participating institutions of the Andean ROAD will share research,
educational and outreach visions and activities.
We expect to increase the engagement with astronomy in all the Andean
ROAD countries, leveraging the special role of Chile with its great
conditions for astronomy and the involvement of many professional
observatories, universities and astronomical institutions. 

\section{Overview of Task Forces}

The Andean ROAD is under supervision of the Office of Astronomy for
Development (OAD) of the IAU. 
The main goal of the Andean ROAD is to guarantee and strengthen
effective methods of communication between the representatives and
coordinators of global, regional and local projects implemented in the
Andean countries, through the three Task Forces established by the OAD
: (i) Astronomy for Universities and Research, (ii) Astronomy for
Children and Schools and (iii) Astronomy for the Public. 
The ROAD Main Coordinator is Jaime Forero-Romero, PhD, from the
Universidad de los Andes (Bogot\'a, Colombia). 
The Andean ROAD's plans include several projects and ideas for each
Task Forces, under supervision of a single representative, as follows:  

\textbf{Task Force 1}. \textit{Coordinator: Germ\'an Chaparro, PhD
  (Universidad ECCI, Bogot\'a, Colombia)}. 

\begin{itemize}
\item Andean School on Astronomy/Astrophysics. The main objective is to
organize a school for advanced undergraduate  and graduate
students. The first version of this school was held in Quito (Ecuador)
during December 2014 with great success, and we expect to develop a
new version during 2016, to be held in Lima (Per\'u), as a joint
effort between Per\'u and Chile. 

\item Andean Graduate Program. The main objective is to determine the
feasibility of creating and funding an Andean graduate program. This
could be a great opportunity to provide powerful experiences in
astronomy for those countries without formal programs. 

\item Massive Open Online Courses. The main objective is to create a MOOC
at the introductory undergraduate level and general public. This MOOC
could also have a version for the general public, and can be
considered as a TF3 project as well.
\end{itemize}

\textbf{Task Force 2}. \textit{Coordinator: \'Angela P\'erez (Parque
  Explora, Medell\'in, Colombia)}.
\begin{itemize}
\item Virtual Training Sessions. The main objective is to create virtual
training and activities open to teachers and students, by developing a
virtual platform to hold virtual workshops on information technologies
applied to astronomy and space sciences. 

\item Special teachings. Design teaching material to work with visually
impaired students. The first step will be a comprehensive research on
this kind of material and current experiences. Later on will be
defined the best way to produce and distribute these materials. 

\item Annual TF2 meeting. An annual meeting to gather all the
collaborators in the Task Force, looking for new collaborations and
alliances. The venue will change every year, with a different key
topic. 
\end{itemize}

\textbf{Task Force 3}. \textit{Coordinator: Farid Char (Sociedad Chilena de Astronom\'ia, Chile)}.

\begin{itemize}
\item Development for planetariums. The main objective is to develop
special shows for planetariums in the region, with the aim to
highlighting ancestral traditions and the particular cosmovision from
many communities in South America. 

\item Communicating Astronomy with the Public. The main objective is to
organize periodic meeting to gather all the collaborators in the Task
Force. This should be a very powerful platform for all people involved
in communicating astronomy to the public. The first meeting will take
place at the same time as the international CAP meeting to be held in
Medell\'in  (16-20 May 2016), only a few months before the 
XV Latin American Regional IAU Meeting, in Cartagena de Indias (3-7
October 2016). 
\end{itemize}

\section{The role of Chile and future plans}

Due to the large size of the Chilean professional astronomical
community and its special role in global astronomy the
interaction to coordinate activities in Chile will be done through the
Sociedad Chilena de Astronom\'ia (SOCHIAS).
SOCHIAS has designated specific representatives per Task Force, and a
main coordinator of the Chilean branch. 
The physical address of the Chilean branch (formally Oficina
Nacional de Coordinaci\'on) is located in the Unidad de Astronom\'ia
of the Universidad de Antofagasta. 

We expect that the Andean ROAD  will  improving the overall state of astronomy
development in our region following the vision of the  IAU Strategic
Plan 2010-2020 \cite[International Astronomical Union   (2012)]{IAU2012}.

More information about our future plans, can be accessed through the
official website of the Andean ROAD
\url{http://andean-road.uniandes.edu.co}.  


\begin{thebibliography}{}

\bibitem[International Astronomical Union (2012)]{IAU2012}
{International Astronomical Union} 2012, 
\textit{Astronomy for Development Strategic Plan 2010-2020}.


\end{thebibliography}

\end{document}
